paper2_resistance_mechanisms_mixtures

### 2016-05-16 paper2_resistance_mechanisms_mixtures.Rmd
add todos
adressing minor issues in check
updating paper2

@@

# The development of insecticide resistance with use of insecticide mixtures and sequences : a simple theoretical exploration.

@@

The development of insecticide resistance is a population genetic process.

The development of resistance is expected to be influenced by :

1. starting frequency of resistance alleles
1. selection coefficient
   + selective advatange of resistance in the presence of the insecticide
1. dominance
   + dominance of the resistance allele relative to the susceptible allele
1. exposure
   + proportion of the insect population that are exposed to the insecticide
1. effectiveness
   + proportion of exposed susceptible insects that are killed by the insecticide

@@

*Maybe start by exploring the influence of the parameters on sole use ?*

@@

```{r,echo=FALSE,results='hide'}
  library(resistance)
```
### Some examples showing how changing effectiveness can alter the ranking of insecticide strategies
#### Curtis' Fig 2. inputs : mix2 > adaptive > mix1 > seq
```{r, eval=TRUE, echo=FALSE, fig.height=4}
runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
```

@@

@@

1. look more at GPIRM
1. mention Curtis
1. do I want y axis of resistance frequency plots to be on a log scale ?
1. can I develop explanatory diagrams to describe the input parameters ?

>>>>

```{r,echo=FALSE,results='hide'}
  library(resistance)
```

@@

# Mechanisms by which insecticide resistance can be influenced by strategies of insectcide use : a simple theoretical exploration.

@@

The changing frequency of insecticide resistance is a population genetic process that can be influenced by :

Parameter  | Description
------------------- | -------------------------------------------------------
1. Effectiveness | proportion of exposed insects with the susceptible allele that are killed by the insecticide
2. Exposure | proportion of the insect population that are exposed to the insecticide
3. Dominance | dominance of the resistance allele relative to the susceptible allele
4. Selection Coefficient  | selective advatange of resistance in the presence of the insecticide
5. Frequency  | frequency of resistance alleles within the population

@@

*accessible summary of the model

@@

*Maybe start by exploring the influence of the parameters on sole use *
*could put here or in the discussion

**single insecticide application effect on increase in resistance of increasing each parameter**

Parameter  | effect on resistance | Mechanism
------------------ | ------- | ---------------------------------------------------
1. Effectiveness | faster | increased deaths of susceptibles
2. Exposure | faster | increased deaths of susceptibles
3. Dominance | faster | increased survival of resistants
4. Selection Coefficient | faster | increased survival of resistants
5. Frequency  | faster | less change needed to reach resistance threshold, and rate of change in resistance tends to increase with frequency

*to help explain I could start from a scenario where the attributes of each insecticide are the same*
*and then explore changing each in turn*
*I could then work up to the curtis fig2 plot*

**When all parameters are equal resistance arises slower for sequential than mixture.**
**Effect of changing parameters on relative performance of mixture versus sequential strategy**

Parameter  | relative performance mix versus seq | Mechanism
------------------ | ------- | ---------------------------------------------------
A. All equal | seq > mix | resistance to the two insecticides in a mixture increases at a similar rate and is reached sooner than if the insecticides are applied individually in sequence
1. Effectiveness | faster | ...
2. Exposure | faster | ...
3. Dominance | faster | ...
4. Selection Coefficient | ...
5. Frequency  | faster | ...

### Some examples showing how changing effectiveness can alter the ranking of insecticide strategies
#### Curtis' Fig 2. inputs : mix2 > adaptive > mix1 > seq
```{r, eval=TRUE, echo=FALSE, fig.height=4}
runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
```

@@

## Discussion

(@)  In much of parameter space resistance changes more slowly when insecticides are used in a mixture relative to sequential use. The mechanism explaining this is ...

(@) For a mixture of 2 insecticides with differing effectiveness. Resistance to the insecticide with a greater effectiveness increases faster. The less effective insecticide is 'protected' by the more effective. Resistance to the less effective insecticide increases slowly until resistance to the more effective insecticide reaches a high level.

(@) When the effectiveness of 2 insecticides in a mixture is similar, the more effective insecticide provides little 'protection' to the less effective. In this case resistance to the two insecticides in a mixture increases at a similar rate and is reached sooner than if the insecticides are applied individually in sequence :

```{r, eval=TRUE, echo=FALSE, fig.height=4}
runcurtis_f2( max_gen=500, P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.17 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 0.75 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.23 , correctMixDeployProp = 1 , addCombinedStrategy = FALSE )
```

## References
1. GPIRM
1. Curtis 
1. Mani GS. Evolution of resistance in the presence of two insecticides. Genetics 1985;109(4):761–83.
1. Birget 2015

@@

1. re-read Mani 1985
1. re-read GPIRM
1. mention Curtis but don't depend on it
1. do I want y axis of resistance frequency plots to be on a log scale ?
1. can I develop explanatory diagrams to describe the input parameters ?
1. decide on base starting scenarios and how to modify params from that
1. mention generation time of 1 month, maybe even rescale axes to years ?
1. look at effect of sex linkage
